todo: think about good title

### 2018-12-16 documentation/paper.tex
started documentation

%}}}
%{{{ graphics
\usepackage{graphicx} % Bilder
%}}}
%{{{ math
\usepackage{mathrsfs} % mathcal and mathscr
\usepackage{mathtools, amssymb, amsthm}
\usepackage{bm} % cool bold symbols
%}}}

% a todo command===============================================================
\newcounter{todocounter}
\newcommand{\todo}[2][noisnotdefined]{
 \marginpar{\fcolorbox{black}{yellow}{\footnotesize\textbf{todo}}
 \ifthenelse{\equal{#1}{noisnotdefined}}{}{\textcolor{black}{\newline\tiny #1}}}
 \textbf{\ifthenelse{\equal{#2}{.}}
   {\fcolorbox{blue}{white}{\textcolor{blue}{$\maltese$}}}{{\textcolor{blue}{#2}}}}
 \refstepcounter{todocounter}}

%===============================================================================

\date{}
\title{Project Proposal for Genetic Algorithms and Evolutionary Programming}
\author{Felix Bartel and D\'avid Kerekes}

@@

\section*{Problem Overview}

@@

\section*{Preliminary Results}

We have done some tests so far with a basic evolutionary algorithm \cite{github_repo} where we fixed the size of the neural network to $6$ input nodes, one hidden layer with $7$ nodes and $2$ output nodes.
As crossover we randomly mixed the nodes with belonging weights and biases of two networks and chose the parents via the fitness proportional roulette wheel selection.
Mutation is accomplished by adding Gaussian noise to a subset of all weights and biases.
Furthermore we implemented elitism.

@@

We implemented our own bot to act as an initial trainer for the network. This trainer is easier to defeat and even gives up after a certain amount of touches with the ball, such that normal gameplay will be rewarded.
The simplification gives the evolutionary algorithm the opportunity to start optimizing the neural network in a valid direction, without introducing additional factors to the fitness score.

\begin{figure}[H]
\center
\includegraphics[width=0.5\textwidth]{img/fitness.png}
\caption{Fitness density of an example run with indicated minimum, average and maximum.}
\label{fig:train}
\end{figure}

The results are depicted in Figure \ref{fig:train}, where the fitness is the normalized point difference after $21$ serves.
One can see that, although we we made simplifications by weakening the opponent, the algorithm produces a network which wins all rounds.

\section*{Further Experiments}

We plan on testing multiple types of crossover and mutation operators, and selection schemes as seen in \cite{montana1989training}. Example mutations include biased and unbiased weight mutations and node mutations. Example crossovers are weight, node, and feature crossovers.

To have a more objective representation of the capabilities of the networks, we plan on reintroducing the original easy, medium and hard difficulty bots in a curriculum-learning fashion.

Finally, as time allows we could extend this work by e.g.\,making the size of the network variable or do self-training with similar techniques as in the paper of Karl Sims\cite{sims1994evolving} presented in the lecture.

>>>>

%}}}
%{{{ graphics
\usepackage{graphicx} % Bilder
%}}}
%{{{ math
\usepackage{mathrsfs} % mathcal and mathscr
\usepackage{mathtools, amssymb, amsthm}
\usepackage{bm} % cool bold symbols
%}}}
%{{{ a todo command
\newcounter{todocounter}
\newcommand{\todo}[2][noisnotdefined]{
 \marginpar{\fcolorbox{black}{yellow}{\footnotesize\textbf{todo}}
 \ifthenelse{\equal{#1}{noisnotdefined}}{}{\textcolor{black}{\newline\tiny #1}} }
 \textbf{\ifthenelse{\equal{#2}{.}}
   {\fcolorbox{blue}{white}{\textcolor{blue}{$\maltese$} } }{ {\textcolor{blue}{#2} } } }
 \refstepcounter{todocounter} }
%}}}

\date{}
\title{todo: think about good title}
\author{Felix Bartel and D\'avid Kerekes}

@@

\section{Introduction}

@@

@@

\section{Standard genetic algorithm}

fixed the size of the neural network to $6$ input nodes, one hidden layer with $7$ nodes and $2$ output nodes.

\subsection{Crossover}

As crossover we randomly mixed the nodes with belonging weights and biases of two networks and chose the parents via the fitness proportional roulette wheel selection.

\subsection{Mutation}

Mutation is accomplished by adding Gaussian noise to a subset of all weights and biases.

\subsection{scheme}

Furthermore we implemented elitism.

fitness is the normalized point difference after $21$ serves.

We implemented our own bot to act as an initial trainer \texttt{trainer} for the network.
This trainer is easier to defeat.
We made a copy of each bot where it gives up after a certain amount of touches with the ball, such that normal gameplay will be rewarded.
The simplification gives the evolutionary algorithm the opportunity to start optimizing the neural network in a valid direction, without introducing additional factors to the fitness score.

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:standard}
\end{figure}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:standard_5t}
\end{figure}
